Molecular Gas Kinematics of the Strongly-lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-11-18 WRITEUP/ms.tex
clean up comments regarding fate of recoil BH

Alternatively, the observed offset %between the BLR and the systemic redshift (and the NLR)
may be explained by a more extreme scenario of a recoiling black hole (BH), 
where its BLR is moving at high velocity relative to the bulk of its host galaxy
\citep{Madau04a, Bonning07a, Loeb07a}.
% where
% BLRs are gravitationally bound to
% the AGN (with a BLR dynamical timescale $\sim$10\E{2} years \citep{Bonning07a}, \ie can be quickly regenerated), 
% whereas this condition is not necessary for NLRs.
%
% In hierarchical model of galaxy formation, supermassive black holes are expected to form through
% mergers, of which the majority took place at high redshifts. 
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a non-zero net angular momentum is being carried away.
Depending on their initial conditions (e.g., black hole mass ratio, spin-orbit orientation, spin magnitude), 
numerical relativity simulations have shown that the recoil velocity can reach up to 
$v_{\rm kick}$\ssim4000\,\kms for spinning BHs, % \citep[\eg][]{Campanelli07a}, 
with a typical offset recoil velocity of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\citep[\eg][]{Libeskind06a,Campanelli07a}.
% ------------- fate ----------------
% Several outcomes of the recoiling blackholes include,
% a) In spherical bulges the black hole will oscillate right through the center, and return to the center after a few dynamical times, or 
% deplete or "heat up" the stars in the nucleus, reducing the efficiency of dynamical and extending the return time.
% b) If the black hole is able to leave the bulge / ejected and enters a triaxial dark matter halo, it may not return at all within a Hubble time \citep{Guedes09a} and become intergalactic/naked black holes (but rare unless in recoil velocity $>>$ few thousands \kms, i.e., sufficient to escape the host galactic potential) \citep[\eg][]{Loeb07a}.
%c) High resolution simulations of gas mergers (e.g. Mayer et al. 2007) show that up to 60\% of the initial gas is funneled to the center of the remnant. In that particular simulation, the mass of the nuclear disk is Md = 3e9 Msun, and recoliing black holes with velocities of 
%$~$500 km/s are not able to escape the central region (scale length L=75 pc).
% In the case of gas-rich merger remnants, where large amounts of gas have been
% funneled to the central regions find that the MBHs remain within R$<$1\,kpc from the center
% of the host even for high recoil velocities ($v_{\rm kick}$\eq1200\,\kms) due to the compactness of the
% remnant galaxy's nuclear disk" \citep{Mayer07a,Guedes08a}.
% ----------------------------------------
%A recoiling black hole may therefore 
%oscillate around the center of the galaxy as its orbit is decaying due to dynamical friction, 
%or leave its host galaxy if $v_{\rm kick}$$\gg$$v_{\rm esc}$ and dynamical friction is inefficient 
%\citep[see \eg \citealt{Volonteri10a} for a discussion]{Guedes09a,Bonning07a, Colpi14a}. 
%% escape velocity
%If we assume a \citet{Hernquist90a} profile for the stellar bulge in RXJ1131, 
%%$M_{\rm BH}$/\mstar\ssim1/375, $v_{\rm circ}$\ssim$v_{\rm rot}, 
%% then the upper limit of $v_{\rm kick}$\ssim2$\times$$v_{\rm circ}$, 
%% bulge escape velocity $v_{\rm esc}$\eq2.83$\v_{\rm circ}$ = 858\kms
%% %(in the case of a density profile of the spherical \citet{Hernquist90a} model for the stellar bulge, 
%% %spherically symmetric isotropic model)%  \citep{BoylanKolchin04a}, 
%then the escape velocity $v_{\rm esc}$\eq2$\sqrt{2}$$v_{\rm circ}$, where $v_{\rm circ}$ is the velocity
%at $r$\eq$a$, the stellar bulge scale length that is at most a few kpc.
%Assuming that the rotation curve of RXJ1131 is similar to that MW, then at inner radius ($\lesssim$ few kpc), 
%we would expect $v_{\rm rot} (r=a)\gtrsim$$v_{\rm rot}$ of the extended gas disk (see \Sec{dynamics}).
%Given that we don't know precisely the stellar bulge scale length, if we assume 
%$v_{\rm rot} (r=a)$\ssim$v_{\rm rot}$ of the gas disk, then $v_{\rm esc}$\ssim860\,\kms. 
%Therefore, the recoil BH in RXJ1131 is unlikely to escape its gas-rich host galaxy.%  \citep[\eg][]{Loeb07a, Volonteri10a}.
Several sources have been proposed as recoiling BH candidates, such as CID-42 at $z$\,=\,0.359 \citep{Civano10a},
SDSS 0956+5127 at $z$\,$\sim$\,0.714 \citep{Steinhardt12a}, and
SDSSJ0927+2943 at $z$\,$\sim$\,0.713, where a set of redshifted narrow lines and a second set of blueshifted broad AGN lines
have been observed \citep{Komossa08a}.
However, this scenario for SDSSJ0927+2943 was recently refuted by \citet{Decarli14a}, who find that the
BLR lines in SDSSJ0927+2943 are indeed consistent with the CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm the redshifted BLR.
Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of previous merger, which is not implausible and 
might also explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

>>>>

Alternatively, the observed offset 
may be explained by a more extreme scenario of a recoiling black hole (BH), 
where its BLR is moving at high velocity relative to the bulk of its host galaxy
\citep{Madau04a, Bonning07a, Loeb07a}.
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a non-zero net angular momentum is being carried away.
Depending on their initial conditions (e.g., black hole mass ratio, spin-orbit orientation, spin magnitude), 
numerical relativity simulations have shown that the recoil velocity can reach up to 
$v_{\rm kick}$\ssim4000\,\kms for spinning BHs, 
with a typical offset recoil velocity of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\citep[\eg][]{Libeskind06a,Campanelli07a}.
Several sources have been proposed as recoiling BH candidates, such as CID-42 at $z$\,=\,0.359 \citep{Civano10a},
SDSS 0956+5127 at $z$\,$\sim$\,0.714 \citep{Steinhardt12a}, and
SDSSJ0927+2943 at $z$\,$\sim$\,0.713, where a set of redshifted narrow lines and a second set of blueshifted broad AGN lines
have been observed \citep{Komossa08a}.
However, this scenario for SDSSJ0927+2943 was recently refuted by \citet{Decarli14a}, who find that the
BLR lines in SDSSJ0927+2943 are indeed consistent with the CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm the redshifted BLR.
Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of previous merger, which is not implausible and 
might also explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.
